---
title: "POLSCI 630 - Problem Set 4"
author: "Huong Vu"
date: "1/31/2022"
output: pdf_document
---
Before loading the data, I loaded the packages: `tidyverse`, `jtools`, `npreg` and `Metrics`. I also set the work directory for the file by going to the tab Session and then Set Working Directory. I loaded the `contrib.csv` dataset which I saved in the same folder with the working directory and named it `contrib`.

```{r setup, results="hide", results="hide", message=F, warning=F}
# load the packages
library(tidyverse)
library(jtools)
library(npreg)
library(Metrics)


contrib <- read_csv("contrib.csv") # load the data
summary(contrib) # quick glimpses at the data
```

# 1. Calculating Standard Errors
The regession model of `given` on `age`:
$$given_i = \beta_0 + \beta_1 \ age_i +  u_i$$
I set Y is `given` which is the dependent variable and X is `age` which is the independent variable. 

```{r, echo=T, results="hide", message=F, warning=F}
# Dependent variable as y
y <- contrib$given
# Independent variable as x
x <- contrib$age
```

I calculated $\beta_0$, $\beta_1$, standard error of $\beta_0$, and standard error of $\beta_1$.

```{r}
# Beta1
beta1 <- sum((y - mean(y)) * (x - mean(x))) / sum(((x - mean(x))^2))
beta1

# Beta0
beta0 <- mean(y) - beta1 * mean(x)
beta0

# Y hat
y_hat <- beta1 * x + beta0

# n
n <- length(x)

# MSE
mse <- (1/(n - 2)) * sum((y - y_hat)^2)
mse

# Standard error of beta1
# Variance of beta1
beta1_var <- mse / sum((x - mean(x))^2)
beta1_var

# Standard error of beta1
beta1_se <- sqrt(beta1_var)
beta1_se

# Standard error of beta0
# Variance of beta0
beta0_var <- mse * ((1/n) + (mean(x)^2/ sum((x - mean(x))^2)))
beta0_var

# Standard error of beta0
beta0_se <- sqrt(beta0_var)
beta0_se

```

I verified the coefficients and the standard errors that I derived by hand using the `lm()` command and reported the results with `summ()`.

```{r pressure, echo=FALSE}
# Compare results to those from the built-in function
agegiven_reg <- lm(given ~ age, data = contrib)
summ(agegiven_reg )
```

# 2. Interpretation and Analysis

* Interpret the $\beta_0$  and $\beta_1$  coefficients: For one unit increase in `age`, it will increase 1.984817 in `given`. It means that if an observation increases 1 year in age, he/she tends to contribute more 1.984817 in `given`, which is equivalent to $1,984.817.

* How much does the model predict that a 50 year old person will give? given_50 = 251.469, equivalent to $251,469
```{r}
given_50 = beta0 + beta1*50
given_50
```

* How much more than a 40-year old will a 60-year old contribute, according to our model? The difference is `39.69634`, equivalent to `39,696.34`.

```{r}
given_40 = beta0 + beta1*40
given_60 = beta0 + beta1*60
given_difference = given_60 - given_40
given_difference
```


* Create a scatter plot depicting the relationship between age and campaign contributions:

```{r}
plot(
contrib$age, contrib$given,
main = "The 2018 Cooperative Congressional Election Survey",
xlab = "Age",
ylab = "Campaign contributions")
abline(agegiven_reg, col = "red") # Add regression object immediately below plot()
```

# 3. Factors Affecting Standard Errors

I extracted a sample of 5,000 observations from the population 500 times and named the dataset of these 5,000 observations `sample_5000`. I performed a linear regression for `sample_5000`. The standard error of the age coefficient is 0.41.

```{r}
set.seed(630) # the set.seed command to ensure that others who run your code will get the same results

for(i in 1:500) { # This code will "loop" 500 times
# Sample 5,000 voters
sample_5000 <- contrib %>%
  sample_n(5000)
# Run regression
sample_5000_reg <- lm(given ~ age, data = sample_5000)
}

summ(sample_5000_reg)


```


I extracted a sample of 2,000 observations from the population 500 times and named the dataset of these 2,000 observations `sample_2000`. I performed a linear regression for `sample_2000`. The standard error of the age coefficient is 0.71

```{r}

set.seed(630) # the set.seed command to ensure that others who run your code will get the same results

for(i in 1:500) { # This code will "loop" 500 times
# Sample 2,000 voters
sample_2000 <- contrib %>%
  sample_n(2000)
# Run regression
sample_2000_reg <- lm(given ~ age, data = sample_2000)
}
summ(sample_2000_reg)

```


Between the two standard errors calculated above, the standard error of the age coefficient from `sample_5000` is smaller than the standard error of the age coefficient from `sample_2000` because the number of observations in `sample_5000` is bigger than in `sample_2000`.  


For `sample_5000`, I filtered out people who are younger than 40 years old and older than 70 years old. The new restricted dataset was named `sample_5000_restricted`. After being filtered out `sample_5000_restricted` only contains 2,791 observations. I performed random draws of 2,000 observations for 500 times and created a linear regression for `sample_5000_restricted`. The standard error of the age coefficient is 1.35.

```{r}
sample_5000_restricted = filter(sample_5000, between(age, 40, 70))
#View(sample_5000_restricted)

set.seed(630) # the set.seed command to ensure that others who run your code will get the same results

for(i in 1:500) {
sam <- sample_5000_restricted %>%
  sample_n(2000)
sam_reg <- lm(given ~ age, data = sam)
}


summ(sam_reg)
```

Between `sample_5000` and `sample_5000_restricted`, the standard error of the age coefficient of `sample_5000` is smaller because the number of observations in `sample_5000_restricted `is smaller.
