---
title: "PUPOLSCI 630 - Problem Set 11"
author: "Huong Vu"
date: "3/29/2022"
output: pdf_document
---

```{r, echo=T, results="hide", message=F, warning=F}
# Libraries

library(tidyverse)
library(haven)
library(jtools)
library(margins)
library(memisc)
library(skedastic)
library(ggplot2)
#install.packages("ggfortify")
library(ggfortify)

# install.packages("remotes")
library(remotes)
install_github("vqv/ggbiplot")
library(ggbiplot)

library(AER)
library(miceadds)
```

# 1.Princiapl Component Analysis (PCA)

```{r, include=FALSE}
dat = read_csv("pca.csv") # Load data
summary(dat) #check NAs
dat <- dat %>% remove_rownames %>% column_to_rownames(var = "country_code") # moving the first column to row names
df <- na.omit(dat)
summary(df)
var(df)
```

## 1.1. Perform principal component analysis with the four variables: dem_polity, fh_pr, dem_vdem, and elec_access. Report the principal component loadings for all variables.

```{r cars}
PCA <- prcomp(df, scale = TRUE) # performing PCA
PCA$rotation # Get the principal component loadings
```

## 1.2. Based on the principal component loadings, what do you think the first principal component is measuring? What do you think the second principal component is measuring? 

From the first principal component, we see the similar weight on `dem_polity`, `dem_vdem` and `fh_pr` and less weight on `elec_access`. So the first component may be the measurement of democracy. And the second component may be the measurement of electricity access.

## 1.3. Report the proportion of variance explained by each principal component. You can calculate it by hand or use a built-in function.

```{r}
summary(PCA) # calculate the proportion of variance explained by each principal component using a built-in function
```

## 1.4. Report the principal component scores for all observations (excluding observations with NA(s) in one or more of the variables).

```{r}
PCA$x
```

## 1.5. Provide a plot of first principal component vs. second principal component.

```{r}
#autoplot(PCA)
autoplot(PCA, data = df, colour = 'elec_access')
```

```{r}
ggbiplot(PCA, labels = rownames(df))
```


# 2. Heteroskedasticity

```{r pressure, echo=FALSE}
schools = read_csv("schools.csv") # Load data
summary(schools)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## 2.1. Regress `story4` on `lunch`, `enroll`, `exppp`, and the interaction between `lunch` and `exppp`. What is the impact of expenditures per pupil on students passing 4th grade reading test when `lunch` = 0 and 30?

The regression model:
$$
story4 = \beta_0 + \beta_1. lunch + \beta_2.enroll + \beta_3.exppp +  \beta_4. lunch\times exppp + u
$$

```{r}
lm <- lm(story4 ~ lunch + enroll + exppp + lunch:exppp, schools)
summ(lm)
```

## 2.2. Perform regression diagnostics using plot(). Does the model seem to suffer from heteroskedasticity?
Looking at the plot, I think that this model seems to suffer from heteroskedasticity

```{r}
plot(lm)
```

## 2.3. Formally test for heteroskedasticity with White test by hand. Interpret your test statistics.

```{r}
y <- schools$story4
x1 <- schools$lunch
x2 <- schools$enroll
x3 <- schools$exppp

```

Test for heteroskedasticity with White test by hand:

```{r}
# Step 1
m_w <- lm(y ~ x1 + x2 + x3 + x1:x3)
squared_residuals <- resid(m_w) ^ 2
predicted <- predict(m_w)
squared_predicted <- predict(m_w) ^ 2

# Step 2
m_w_stage2 <- lm(squared_residuals ~ predicted + squared_predicted)
R_squared_w <- summary(m_w_stage2)$r.squared
n <- length(y) # N observations

# Step 3
(lm_st_w <- n*(R_squared_w))
(1 - pchisq(lm_st_w, df = 2))
```

## 2.4. Formally test for heteroskedasticity with the Breusch-Pagan test and the White test using R’s built-in functions. Interpret your test statistics.

```{r}
bptest(lm)
```

* The test statistic is X^2 = 131.88
* The degrees of freedom is 4
* The corresponding p-value is 2.2e-16
* Ho: Homoscedasticity is present
* H1: heteroskedasticity is present
* We can reject the null because p-value is very small --> This model suffers from heteroskedasticity

```{r}
y_hat_schools <- predict(m_w)
bptest(m_w, varformula = ~ y_hat_schools + I(y_hat_schools^2))
```

* The test statistic is X^2 = 157.28
* The degrees of freedom is 2
* The corresponding p-value is 2.2e-16
* Ho: Homoscedasticity is present
* H1: heteroskedasticity is present
* We can reject the null because p-value is very small --> This model suffers from heteroskedasticity


## 2.5. Suppose you suspect that heteroskedasticity is present and decide to use robust standard errors to be safe. Calculate robust standard errors for your model by hand. Verify with a built-in function.

\[
\hat{Var} (\hat \beta_j) = \frac{\sum_{i=1}^{n} \hat r^2_{ij}\hat u_i^2}{SSR_j^2}
\]

```{r}
m_robust1 <- lm(x1 ~ x2 + x3 + x1:x3)
m_robust2 <- lm(x2 ~ x1 + x3 + x1:x3)
m_robust3 <- lm(x3 ~ x1 + x2 + x1:x3)
m_robust4 <- lm(I(x1*x3) ~ x1 + x2 + x3)
```

```{r}
numerator1 <- sum((resid(m_robust1)**2) * (resid(lm)**2))
SSR1 <- sum(resid(m_robust1)**2) ** 2
(var_beta_1 <- numerator1 / SSR1) # variance
```

```{r}
(robust_se_beta_1 <- sqrt(var_beta_1)) # robust std. error
```

```{r}
numerator2 <- sum((resid(m_robust2)**2) * (resid(lm)**2))
SSR2 <- sum(resid(m_robust2)**2) ** 2
(var_beta_2 <- numerator2 / SSR2)
```

```{r}
(robust_se_beta_2 <- sqrt(var_beta_2)) # robust std. error
```

Verify by built-in function

```{r}
var <- vcovHC(lm, type = "HC") # from the sandwich package (through AER)
var
```

```{r}
robust_se <- sqrt(diag(var))
robust_se
```

