---
title: "POLSCI 630 - Problem Set 3"
author: "Huong Vu"
date: "1/22/2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

Before loading the data, I loaded the `tidyverse` and `jtools`package. I also set the work directory for the file by going to the tab Session and then Set Working Directory. I loaded the `ceosalary.csv` dataset which I saved in the same folder with the working directory and named it `ceosalary`.

```{r, echo=T, results="hide", message=F, warning=F}
# set universal settings for all chunks
knitr::opts_chunk$set(warning = F, message = F,      # don't print warnings or messages. 
                      fig.width = 5, fig.height = 4) # format all figures as 5x4 inches.

# load the packages
library(tidyverse) # basic package for data wrangling functions
library(jtools)
ceosalary <- read_csv("ceosalary.csv") # load the data
```

# 1. Bivariate Linear Regression

I set Y is `salary` which is the dependent variable and X is `profits` which is the independent variable. I calculated the values of $\beta_0$, $\beta_1$, TSS, ESS, SSR, $R^{2}$ and got the results as follows:

* The slop of the equation is equal to $\beta_1$ = 0.5722961
* The intercept of the equation is equal to $\beta_0$ = 746.92381

* A CEO who made \$0 in firm profits is predicted to make $\beta_0$ = 746.92381 (intercept), equivalent to \$746,923.81 in salary in 1990.
* A CEO who made \$10-million in firm profits is predicted to make predicted $\hat{y}$ = 752.6468, equivalent to \$752,646.8 in salary in 1990 (this calculation is conducted in Part 2 of this Problem Set).
* A \$1-million increase in firm profits is associated with an increase of $\beta_1$ = 0.5722961, equivalent to \$572.2961 in the salary of the CEO.

```{r}

# y as dependent variable
y <- ceosalary$salary
# x as independent variable
x <- ceosalary$profits

beta1 <- sum((x - mean(x)) * (y - mean(y))) / sum(((x - mean(x))^2)) # calculate beta 1
beta1

beta0 <- mean(y) - beta1 * mean(x) # calculate beta 0
beta0

# predicted y hat
y_hat <- beta0 + beta1 * x 
y_hat

# Compare predicted to actual values 
ceosalary$salary_hat <- y_hat
ceosalary$u2 <- (y - y_hat)^2
#View(ceosalary)

# Run regression with built-in function
ceosalary_reg <- lm(salary ~ profits, data = ceosalary)
summary(ceosalary_reg)

# a function from the jtools package that allows us to view the results from regression
summ(ceosalary_reg)

# total sum of squares
tss <- sum((y - mean(y))^2)
tss

# explained sum of squares
ess <- sum((y_hat - mean(y))^2)
ess

# sum squares of errors
ssr <- sum(((y - y_hat)^2 ))
ssr

# r squared
r_2 = ess/tss
r_2

```

# 2. Creating a Scatter Plot

The scatter plot depicting the relationship between salary and profits: the x-axis demonstrates the values of the variable `profits` and the y-axis demonstrates the values of the `salary` variable. The regression line is in blue.

```{r}
# plot 1
plot(
  ceosalary$profits, ceosalary$salary,
  main = "CEO Salary",
  xlab = "Profits",
  ylab = "Salary")
abline(ceosalary_reg, col = "blue") # Add regression object immediately below plot()
```

The scatter plot predicting salary when profit is $10 million: the x-axis demonstrates the values of the variable `profits` and the y-axis demonstrates the values of the `salary` variable. The regression line is in blue. The point is located by the red vertical and horizontal `abline`.

```{r}
# plot 2
x_val1 <- 10
x_val1

# predicted y value
beta1 * x_val1 + beta0

beta0


plot(
  ceosalary$profits, ceosalary$salary, 
  main = "CEO Salary", 
  xlab = "Profits", 
  ylab = "Salary"
  )
abline(ceosalary_reg, col = "blue")
abline(v = x_val1, col = "red")
abline(h = (beta1 * x_val1 + beta0), col = "red")
```

# 3. Using for Loop

I created a new column in the dataset named `salaryeuro` that reports the salary in €1000 (Euro) with the conversion rate $1 = €0.88.

```{r}
# use `for` loop to create new variable `salaryeuro`
#mutate(ceosalary, salaryeuro = salary * 0.88)
ceosalary$salaryeuro <- 0
for (i in 1:length(ceosalary$salaryeuro)) { ceosalary$salaryeuro[i] <- ceosalary$salary[i] * 0.88 } 
#View(ceosalary)
```

I used a for loop to count the number of CEOs who are older than 60 years old. There are a total of 53 CEOs who are older than 60.

```{r}

count <- 0
for (i in ceosalary$age) {
  if (i > 60) {
    count = count + 1}
  }

print(count)
```

I created a new variable named `profitsbinary`. The `profitsbinary` variable will take a value of 1 if a CEO’s firm made a positive profit (`profits` > 0) and a value of 0 if the CEO’s firm suffered a loss (`profits` < 0).

```{r}
# create a new variable `profitsbinary`
ceosalary$profitsbinary <- 0 

# create for loop to assign the values for `profitsbinary`
for (i in 1:length(ceosalary$profits)) {
  if (ceosalary$profits[i] > 0 ) {
    ceosalary$profitsbinary[i] <- 1
  }else {
      ceosalary$profitsbinary[i] <- 0
  }
}

#View(ceosalary)
#It run successfully

```









